Parsing Postfix log files

### 2009-03-26 doc/logparser-implementation.tex
More corrections from proofreading.

XXX WRITE AN INTRODUCTION\@.  MOST OF THIS IS CRAP, BUT I MIGHT BE ABLE TO
RESCUE SOME OF IT\@.

@@

    \item [DISCONNECT] Invoked when the remote client disconnects, it
        enters the connection in the database, performs any required
        cleanup, and deletes the connection from the state tables.  This
        action deals with aborted delivery attempts (\sectionref{aborted
        delivery attempts}).

    \item [EXPIRY] If Postfix has not managed to deliver a mail after
        trying for five days it will give up and return the mail to the
        sender.  When this happens the mail will not have a combination of
        Postfix programs which passes the valid combinations check as
        described in \sectionref{out of order log lines}; this action, with
        cooperation from the COMMIT and SAVE\_DATA actions, deals with that
        complication.

    \item [MAIL\_BOUNCED] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to help distinguish delivery attempts that bounce from other
        delivery attempts.

@@

    \item [MAIL\_SENT] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to help distinguish successful delivery attempts from other
        delivery attempts.

    \item [MAIL\_TOO\_LARGE] When a client tries to send a mail larger than
        the local server accepts it will be discarded and the client
        informed of the problem.  The mail may have been accepted and
        partially transferred, in which case there's a data structure to
        dispose of, otherwise there will not be.  See \sectionref{timeouts
        during data phase} for the gory details; although that describes
        timeouts the processing is the same for mails that are too large.

    \item [PICKUP] The PICKUP action corresponds to the \daemon{pickup}
        service dealing with a locally submitted mail.  A new mail will be
        created, although out of order log entries may have caused the
        state table entry to already exist, as documented in
        \sectionref{pickup logging after cleanup}.

    \item [POSTFIX\_RELOAD] When Postfix stops or reloads its configuration
        it kills all \daemon{smtpd} processes, requiring any active
        connections to be cleaned up, entered in the database, and deleted
        from the state tables.  Postfix probably kills all the other
        components too, but \parsername{} is only affected by \daemon{smtpd}
        processes exiting.

@@

    \item [SMTPD\_DIED] Sometimes a \daemon{smtpd} dies or exits
        unsuccessfully; the active connection for that \daemon{smtpd} must
        be cleaned up, entered in the database if it has sufficient data,
        and deleted from the state tables.  In some cases the connection
        will not have enough data to satisfy the database schema, so it
        cannot be entered into the database; unfortunately this means that
        some data is lost.

    \item [SMTPD\_KILLED] Sometimes a \daemon{smtpd} is killed by a signal;
        this is processed in the same way as a \daemon{smtpd} dying, see
        the \texttt{SMTPD\_DIED} action for details.

@@

    \item [TIMEOUT] The connection with the remote client timed out so the
        mail currently being transferred must be discarded.  The mail may
        have been accepted, in which case there's a data structure to
        dispose of, or it may not in which case there is not.  See
        \sectionref{timeouts during data phase} for full details.

@@

The rules are responsible for classifying each log line and specifying the
correct action to be invoked.  For any parser implemented using this
architecture, the rules are the most visible component and most likely to
be modified by users.  Rules need to be as simple as possible so that users
can easily modify or add to them, but the implementation must balance that
simplicity with the need to provide enough flexibility and power to
successfully parse the inputs.

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that abstract description will not be repeated
here, but should be fleshed out by the concrete example rule in
\sectionref{example rule in implementation}.  The process of creating new
rules from unparsed log lines is dealt with in \sectionref{creating new
rules in implementation}; the implementation of the utility supplied with
\parsername{} to create regexes from unparsed log lines is also described
in that section.  The rule conditions used in \parsername{} are explained
next, and the final portion of this section provides some
suggestions for how to detect overlapping rules.

@@

The example rule in \tableref{Example rule in implementation table}
recognises the log line resulting from Postfix rejecting a delivery attempt
because the domain in the sender address does not have an A, AAAA,
or MX DNS entry, i.e.\ mail could not be delivered to the sender's address
(for full details see~\cite{reject-unknown-sender-domain}).

@@

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and regex.  The name and description fields should be
set based on the meaning of the log line, to help others understand which
log lines this rule will recognise; the program will be obvious from the
unrecognised log lines.  The action depends on the what the log line
represents, e.g.\ a delivery rejection, a mail being delivered, some useful
information, or something else; examine the list of actions in
\sectionref{actions in detail in implementation} to determine the correct
one.  The regex will be constructed based on the log line, but see below
for a tool to ease the process.

There are other attributes that may be required in a rule:
connection\_data, result\_data, priority, or restriction\_name.  In general
it will only become clear that connection\_data or result\_data are
required when \parsername{} warns about an entry in the connections or
results tables that is missing some required fields, because values for the
missing fields are not present in any of the log lines.  E.g.\ the rule
recognising the \daemon{pickup} component processing a mail sets the
client\_hostname to \texttt{localhost} and the client\_ip to
\texttt{127.0.0.1}, because the mail originates on the local server.  If
the new rule deliberately uses the architecture's overlapping rules feature
it will be obvious that the priority field needs to be set, on this rule and
possibly others; the priority field may be needed on unintentionally
overlapping rules too, but that is more difficult to determine.  Finally
the restriction\_name field should be set when the rule's action is
\texttt{DELIVERY\_REJECTED}; hopefully the restriction in question will be
clear from the content of the log line.

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: log lines are generally created by substituting variable words into
a fixed pattern, and analysis of the frequency with which each word occurs
can be used to determine whether the word is variable or part of the fixed
pattern.  This classification can be used to group similar log lines and
generate a regex to match each group of log lines.  There are 4 steps in
the algorithm:

@@

    \item [Pre-process the file.]  The new algorithm leverages the
        knowledge gained when writing rules and developing \parsername{};
        it performs a large number of substitutions on the input log lines,
        replacing commonly occurring variable terms (e.g.\ email addresses,
        \acronym{IP} addresses, the standard start of rejection messages)
        with keywords that \parsername{} will expand when it loads the new
        rule.  The purpose of this step is to utilise existing knowledge to
        create more accurate regexes; it replaces a large number of
        variable words with fixed words, improving the subsequent
        classification of words as fixed or variable.  The altered log
        lines are written to a temporary file, which subsequent stages will
        use instead of the original input file.

@@

    \item [Calculate word frequencies.]  The position of words within a log
        line is important: a word occurring in two log lines does not
        indicate similarity unless it occupies the same position in both
        log lines.  If a variable term within a line contains
        spaces, it will appear to the algorithm as more than one word.
        This will alter the position of subsequent words, so a word
        occurring in different positions in two log lines \textit{may\/}
        indicate similarity, but the algorithm does not attempt to deal
        with this possibility.  The algorithm maintains a counter for each
        \texttt{(word, word's position within the log line)} tuple,
        incrementing it each time that word occurs in that position.

        The original algorithm hashes each word and checks that result's
        counter from the previous pass; if that counter has a high
        frequency a separate counter will be maintained for this word in
        this position.  This reduces the number of counters maintained by
        the algorithm in this step, reducing the memory requirements of the
        algorithm at a cost of increased CPU usage.

        As time goes on the amount of memory typically available increases
        and the need to reduce memory requirements decreases, so the
        modified algorithm omits the hashing step and maintains counters
        for all tokens.  In addition most of the infrequently occurring
        words will have been substituted with keywords during the first
        step, vastly reducing the number of tuples to maintain counters
        for.

    \item [Classify words based on their frequency.]  The frequency of
        every tuple \texttt{(word, word's position within the log line)} is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new regexes.  Variable terms are
        replaced by \texttt{.+} to match one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly reassembled
        regexes are collapsed to a single \texttt{.+}; any resulting
        duplicate regexes are combined, and their counters added together.
        If a regex's counter is lower than the threshold supplied by the
        user the regex is discarded; this second threshold is independent
        of the threshold used to differentiate between fixed and variable
        words, but once again 1\% of log lines is a good starting point.
        The new regexes are displayed for the user to add to the database,
        either as new rules or merged into the regexes of existing rules;
        the counter for each regex is also displayed, giving the user an
        indication of how many of the log lines that regex should match.
        Discarding regexes will result in some of the log lines not being
        matched; when the rules have been augmented with the new regexes,
        the original log files should be parsed again, and any remaining
        unparsed log lines used as input to this utility.

@@

XXX MERGE logs2regexes WITH check-regexes.

@@

Rule conditions as part of the architecture have already been documented in
\sectionref{rule conditions in architecture}, and they can be very complex
and difficult to evaluate.  In contrast, \parsername{} uses very simple
rule conditions: each rule has a \texttt{program} attribute that specifies
the Postfix component whose log lines it recognises, and each log line
contains the name of the Postfix component that produced it; the framework
will only use rules where the two are equal when trying to recognise log
lines.  This avoids needlessly trying rules that will not match the log
line, or worse, might match unintentionally.  In addition the framework
supports generic rules, whose program attribute is ``*''; these will be
used if none of the program-specific rules recognise the log line.  If none
of the rules are successful the framework will warn the user.

@@

    \item It simplifies both reading and writing of regexes and makes each
        regex largely self-documenting.  E.g.\ the meaning of
        \_\_CLIENT\_HOSTNAME\_\_ is immediately clear, whereas
        \verb!(?:unknown|(?:[-.\w]+))! needs to be deciphered.

    \item It avoids needless repetition of complex regex components, and
        allows the components to be corrected or improved in one
        location.  E.g.\ \_\_SENDER\_\_ is used in 68 rules; if a mistake
        is discovered it can be corrected in one place only.

    \item It enables automatic extraction and saving of captured data.  The
        regex snippets use Perl 5.10's named capture buffers~\cite{perlre},
        so the mapping between captures and fields does not need to be
        explicitly specified by the rule.

@@

    \item [\_\_COMMAND\_\_]  All \acronym{SMTP} commands, except
        \texttt{DATA}, which typically has a more specific rule.
        Priorities could have been used instead of excluding \texttt{DATA}.

@@

    \item [\_\_SHORT\_CMD\_\_]  Postfix sometimes logs \acronym{SMTP}
        commands in a short, single word form.

@@

\parsername{} does not try to detect overlapping rules; that responsibility
is left to the author of the rules.  The advantages and difficulties of
overlapping rules have already been addressed in \sectionref{overlapping
rules in architecture} and will not be repeated here.  \parsername{}
provides a mechanism for ordering overlapping rules: the priority field in
each rule (defaulting to zero); negative priorities may be useful for
catchall rules.

@@

    \item Sort by program and regex, then visually inspect the list, e.g.\
        with \acronym{SQL} similar to:      \newline{}
        \verb!SELECT program, regex!        \newline{}
        \verb!    FROM rules!               \newline{}
        \verb!    ORDER BY program, regex;! \newline{}
        The rules are sorted first by program, then by regex, because rules
        with overlapping regexes will not overlap if their programs are
        different.  Note that this does not properly deal with rules whose
        program is \texttt{*}; those rules will be used on all log lines
        that have not been recognised by program-specific rules.

    \item Compare the results of parsing using different rule orderings
        (described in \sectionref{rule ordering for efficiency}).  Parse
        several log files using optimal ordering, then dump a textual
        representation of the rules, connections, and results tables.
        Repeat with shuffled and reversed ordering, starting with a fresh
        database.  If there are no overlapping rules the tables from each
        run will be identical; differences indicate overlapping rules.  The
        rules that overlap can be determined by examining the differences
        in the tables: each result contains a reference to the rule which
        created it, which will change if that rule overlaps with another.
        Unfortunately this method cannot prove the absence of overlapping
        rules; it can detect overlapping rules, but only if there are log
        lines in the log files that are recognised by more than one rule.

>>>>

XXX WRITE AN INTRODUCTION\@.  MOST OF THIS IS CRAP, BUT I MIGHT BE ABLE TO
RESCUE SOME OF IT\@.  IT SHOULD HAVE A BLURB, FOLLOWED BY A PREVIEW OF THE
CHAPTER\@.

XXX MAIL MADS ABOUT EXAMINERS\@.

@@

    \item [DISCONNECT] Invoked when the remote client disconnects, it
        enters the connection in the database if it has any useful data,
        performs any required cleanup, and deletes the connection from the
        state tables.  This action deals with aborted delivery attempts
        (\sectionref{aborted delivery attempts}).

    \item [EXPIRY] If Postfix has not managed to deliver a mail after
        trying for five days, it will give up and return the mail to the
        sender.  When this happens the mail will not have a combination of
        Postfix programs which passes the valid combinations check,
        implemented to deal with the complication described in
        \sectionref{out of order log lines}; this action tags the mail as
        having expired, so the COMMIT action will skip the valid
        combinations check.

    \item [MAIL\_BOUNCED] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish delivery attempts that bounce from other
        delivery attempts.

@@

    \item [MAIL\_SENT] This action behaves exactly like the
        \texttt{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish successful delivery attempts from other
        delivery attempts.

    \item [MAIL\_TOO\_LARGE] When a client tries to send a mail larger than
        the local server accepts, the mail will be discarded and the client
        informed of the problem.  The mail may have been accepted and
        partially transferred, in which case there's a data structure to
        dispose of.  See \sectionref{timeouts during data phase} for full
        details; although that describes timeouts the processing is the
        same for mails that are too large.

    \item [PICKUP] The PICKUP action corresponds to the \daemon{pickup}
        service dealing with a locally submitted mail.  A new mail will be
        created, although out of order log entries may have caused it to
        already exist, as documented in \sectionref{pickup logging after
        cleanup}.

    \item [POSTFIX\_RELOAD] When Postfix stops or reloads its
        configuration, it kills all \daemon{smtpd} processes, requiring all
        connections to be cleaned up, entered in the database, and deleted
        from the state tables.  Postfix probably kills all the other
        components too, but \parsername{} is only affected by
        \daemon{smtpd} processes exiting.

@@

    \item [SMTPD\_DIED] Sometimes a \daemon{smtpd} dies, is killed by a
        signal, or exits unsuccessfully; the associated connection must be
        cleaned up, entered in the database if it has sufficient data, and
        deleted from the state tables.  In some cases the connection will
        not have enough data to satisfy the database schema, so it cannot
        be entered into the database; unfortunately this means that some
        data is lost.

@@

    \item [TIMEOUT] The connection with the remote client timed out, so the
        mail currently being transferred must be discarded.  The mail may
        have been accepted, in which case there's a data structure to
        dispose of.  See \sectionref{timeouts during data phase} for full
        details.

@@

The rules are responsible for classifying each log line and specifying the
correct action to be invoked.  In any parser implemented using this
architecture, the rules will be the most visible component and most likely
to be modified by users.  Rules need to be as simple as possible so that
users can easily modify or add to them, but the implementation must balance
that simplicity with the need to provide enough flexibility and power to
successfully parse inputs.

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that description will not be repeated here, but
should be fleshed out by the example rule in \sectionref{example rule in
implementation}.  The process of creating new rules from unparsed log lines
is dealt with in \sectionref{creating new rules in implementation}; the
implementation of the utility supplied with \parsername{} to create regexes
from unparsed log lines is also described in that section.  The rule
conditions used in \parsername{} are explained next, and the final portion
of this section provides some suggestions for how to detect overlapping
rules.

@@

The example rule in \tableref{Example rule in implementation table}
recognises the log line resulting from Postfix rejecting a delivery attempt
because the domain in the sender address does not have an A, AAAA, or MX
DNS entry, i.e.\ mail could not be delivered to any address in the sender's
domain (for full details see~\cite{reject-unknown-sender-domain}).

@@

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and regex.  The name and description fields should be
set based on the meaning of the log line, to help others understand which
log lines this rule will recognise; the program will be obvious from the
unrecognised log lines.  The action depends on the what the log line
represents, e.g.\ a delivery rejection, a mail being delivered, some useful
information, or something else; examine the list of actions in
\sectionref{actions in detail in implementation} to determine the correct
one.  The regex needs to be constructed based on the log line, but see
below for a tool to ease the process.

There are other attributes that may be required to make a rule work
correctly: connection\_data, result\_data, priority, or restriction\_name.
In general it will only become clear that connection\_data or result\_data
are required when \parsername{} warns about an entry in the connections or
results tables that is missing some required fields, because values for the
missing fields are not present in any of the log lines.  E.g.\ the rule
recognising the \daemon{pickup} component processing a mail sets the
client\_hostname to \texttt{localhost} and the client\_ip to
\texttt{127.0.0.1}, because the mail originates on the local server.  If
the new rule deliberately uses the architecture's overlapping rules feature
it will be obvious that the priority field needs to be set, on this rule
and possibly others; the priority field may be needed on unintentionally
overlapping rules too, but that is more difficult to determine.  Finally
the restriction\_name field should be set when the rule's action is
\texttt{DELIVERY\_REJECTED}; hopefully the name of the restriction will be
clear from the content of the log line.

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the general
explanation below.  The core of the \acronym{SLCT} algorithm is quite
simple: programs generally create log lines by substituting variable words
into a fixed pattern, and analysis of the frequency with which each word
occurs can be used to determine whether the word is variable or part of the
fixed pattern.  This classification can be used to group similar log lines
and generate a regex to match each group of log lines.  There are 4 steps
in the algorithm:

@@

    \item [Pre-process the file.]  The new algorithm leverages the
        knowledge gained when writing rules and developing \parsername{};
        it performs a large number of substitutions on the input log lines,
        replacing commonly occurring variable terms (e.g.\ email addresses,
        \acronym{IP} addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces a large number of variable words
        with fixed words, improving the subsequent classification of words
        as fixed or variable.  The altered log lines are written to a
        temporary file, which subsequent stages will use instead of the
        original input file.

@@

    \item [Calculate word frequencies.]  The position of words within a log
        line is important: a word occurring in two log lines does not
        indicate similarity unless it occupies the same position in both
        log lines.  If a variable term substituted into a log line contains
        spaces, it will appear to the algorithm as more than one word.
        This will alter the position of subsequent words, so a word
        occurring in different positions in two log lines \textit{may\/}
        indicate similarity, but the algorithm does not attempt to deal
        with this possibility.  The algorithm maintains a counter for each
        \texttt{(word, word's position within the log line)} tuple,
        incrementing it each time that word occurs in that position.

        The original algorithm hashes each word and checks that result's
        counter from the previous pass; if that counter has a high
        frequency a separate counter will be maintained for this
        \texttt{(word, word's position within the log line)} tuple.  This
        reduces the number of counters maintained by the algorithm in this
        step, reducing the memory requirements of the algorithm at the cost
        of increased CPU usage.

        As time goes by the amount of memory typically available to a
        program or algorithm increases, and the need to reduce memory
        requirements correspondingly decreases, so the modified algorithm
        omits the hashing step and maintains counters for all tokens.  Most
        of the infrequently occurring words will have been substituted with
        keywords during the first step, vastly reducing the number of
        tuples to maintain counters for; the original algorithm does not
        have the detailed knowledge leveraged by the new algorithm, because
        it is a generic tool.

    \item [Classify words based on their frequency.]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that will hopefully be
        noticed by the user when reviewing the new regexes, and will be
        obvious when the new rules do not recognise some log lines they are
        expected to.  Variable terms are replaced by \texttt{.+} to match
        one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes are combined, and their counters added together.  If a
        regex's counter is lower than the threshold supplied by the user
        the regex is discarded; this second threshold is independent of the
        threshold used to differentiate between fixed and variable words,
        but once again 1\% of log lines is a good starting point.  The new
        regexes are displayed for the user to add to the database, either
        as new rules or merged into the regexes of existing rules; the
        counter for each regex is also displayed, giving the user an
        indication of how many of the log lines that regex should match.
        Discarding regexes will result in some of the log lines not being
        matched; when the rules have been augmented with the new regexes,
        the original log files should be parsed again, and any remaining
        unparsed log lines used as input to this utility.

@@

XXX MERGE logs2regexes WITH check-regexes.  CHECK IF SLCT DOES THIS STEP\@.

@@

Rule conditions as part of the architecture have already been documented in
\sectionref{rule conditions in architecture}, and they can be very complex
and difficult to evaluate.  In contrast, \parsername{} uses very simple
rule conditions: each rule has a \texttt{program} attribute that specifies
the Postfix component whose log lines it recognises, and each log line
contains the name of the Postfix component that produced it; when trying to
recognise log lines, the framework will only use rules where the two are
equal.  This avoids needlessly trying rules that will not recognise the log
line, or worse, might recognise it unintentionally.  In addition the
framework supports generic rules, whose program attribute is ``*''; these
will be used if none of the program-specific rules recognise the log line.
If none of the rules are successful the framework will warn the user.

@@

    \item It simplifies both reading and writing of regexes and makes each
        regex largely self-documenting.  E.g.\ the meaning of
        \_\_CLIENT\_HOSTNAME\_\_ is immediately clear, whereas
        \verb!(?:unknown|(?:[-.\w]+))! needs to be deciphered each time it
        is encountered.

    \item It avoids needless repetition of complex regex components, and
        allows the components to be corrected or improved in one location.
        E.g.\ \_\_SENDER\_\_ is used in 68 rules; if a mistake is
        discovered it only needs to be corrected in one place.

    \item It enables automatic extraction and saving of captured data.  The
        regex snippets use Perl 5.10's named capture buffers~\cite{perlre}
        to capture data, so the mapping between captures and fields does
        not need to be explicitly specified by the rule.

@@

    \item [\_\_COMMAND\_\_]  All \acronym{SMTP} commands.

@@

    \item [\_\_SHORT\_CMD\_\_]  Postfix sometimes logs \acronym{SMTP}
        commands in a short, single word form; this snippet matches all of
        those, except \texttt{DATA}, which typically has a more specific
        rule.  Priorities could have been used instead of excluding
        \texttt{DATA}.

@@

The advantages and difficulties of overlapping rules have already been
addressed in \sectionref{overlapping rules in architecture} and will not be
repeated here.  \parsername{} does not try to detect overlapping rules;
that responsibility is left to the author of the rules.  A mechanism is
provided for ordering overlapping rules: the priority field in each rule;
negative priorities may be useful for catchall rules.

@@

    \item Sort rules by program and regex, then visually inspect the list,
        e.g.\ with \acronym{SQL} similar to:    \newline{}
        \verb!SELECT program, regex!            \newline{}
        \verb!    FROM rules!                   \newline{}
        \verb!    ORDER BY program, regex;!     \newline{}
        The rules are sorted first by program, then by regex, because rules
        with overlapping regexes will not overlap if their programs are
        different.  Note that this does not properly deal with generic
        rules whose program is \texttt{*}; those rules will be used on all
        log lines that have not been recognised by program-specific rules.

    \item Compare the results of parsing using different rule orderings, as
        described in \sectionref{rule ordering for efficiency}.  Parse
        several log files using optimal ordering, then dump a textual
        representation of the rules, connections, and results tables.
        Repeat with shuffled and reversed ordering, starting with a fresh
        database.  If there are no overlapping rules the tables from each
        run will be identical; differences indicate overlapping rules.
        The rules that overlap can be determined by examining the
        differences in the tables: each result contains a reference to the
        rule which created it, which will change if that rule overlaps with
        another.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the log files that are recognised by more
        than one rule.
